Varying paths to building the social state in the progressive era

### 2023-11-02 t1_analysis.Rmd
modeling beginnings

---
title: "Varying paths to building the social state in the progressive era"
author: "Frank Edwards"
date: "2023-08-15"
output: bookdown::html_document2
---

@@

@@

```{r read, cache = T}
### read employment data
### full ts of occupations ADD CONSTABLES, THERES A DECENT NUMBER OF THEM
dat_occ<-read_csv("./data/usa_00071.csv") %>% 
  rename_all(tolower) %>% 
  mutate(occ1950 = 
           case_when(
             occ1950 == 79 ~ "socwork",
             occ1950 == 773 ~ "police",
             occ1950 == 771 ~ "police",
             occ1950 == 782 ~ "police",
             occ1950 == 93 ~ "teachers"
           )) %>% 
  group_by(year, statefip, occ1950) %>% 
  summarize(emp = sum(perwt)) %>% 
  pivot_wider(names_from = occ1950, values_from = emp,
              values_fill = 0)
### AK OR HI HERE? check that at IPUMS

# pop data (total, age, race, nativitiy, urban)
dat_immig<-read_csv("./data/usa_00070.csv") %>% 
  rename_all(tolower) %>% 
  group_by(year, statefip) %>% 
  summarize(tot_pop = sum(perwt),
            children = sum((age<18) * perwt),
            white_immig = 
              sum((nativity == 5) * (race == 1) * (hispan == 0) * perwt),
            blk = 
              sum((race == 2) * perwt),
            urban = sum(hhwt * (urban ==2)),
            tot_hh = sum(hhwt)) 
### AK OR HI HERE? check that at IPUMS

### group quarters data, 100% 
gq<-read_csv("./data/usa_00069.csv") %>% 
  rename_all(tolower) %>% 
  filter(age<18) %>% 
  group_by(year, statefip, gqtype) %>% 
  summarize(gq_pop = sum(perwt, na.rm = T)) %>% 
  mutate(gqtype = case_when(
    gqtype == 2 ~ "child_inst_correctional",
    gqtype == 3 ~ "child_inst_mental",
    gqtype == 4 ~ "child_inst_poor_disabled")) %>% 
  filter(gqtype!="other") %>% 
  pivot_wider(names_from = gqtype, values_from = gq_pop,
              values_fill = 0)
### AK OR HI HERE? check that at IPUMS

### join
dat <- data.frame(state = state.abb, 
               region = state.region, 
               division = state.division) %>% 
  left_join(state.fips %>% 
              select(fips, abb) %>% 
              distinct() %>% 
              rename(state = abb, statefip = fips)) %>% 
  left_join(dat_immig) %>% 
  left_join(dat_occ) %>% 
  ungroup() %>% 
  left_join(gq)

### homestead act data
#### load in homestead data and boarding school data
### state measure can be cumulative sum of homesteads, that gets at polity
### homesteading data
y_o_s <- read_csv("./data/year_of_statehood.csv") %>% 
  left_join(dat %>% 
              select(state, region) %>% 
              distinct())

hs<-read_csv("./data/blm_homesteads_clean.csv") %>% 
  pivot_longer(claims1868:acres1961,
               names_to = "name",
               values_to = "value") %>% 
  mutate(year = str_sub(
    name, start = -4, end = -1),
    type = str_sub(
      name, start = 1, end = -5)) %>% 
  select(-name) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(value = ifelse(is.na(value),
                        0,
                        value))

### split Dakota Territory values to 0.5 ND, 0.5 SD to preserve time series
hs_dakota<-hs %>% 
  filter(state == "Dakota Territory") %>% 
  mutate(value = value / 2) %>% 
  select(-state)

hs<-hs %>% 
  bind_rows(hs_dakota %>% 
              mutate(state = "North Dakota")) %>% 
  bind_rows(hs_dakota %>% 
              mutate( state = "South Dakota")) %>% 
  filter(state != "Dakota Territory") %>% 
  group_by(state, year, type) %>% 
  summarize(value = sum(value))

hs<-hs %>% 
  arrange(state, year) %>% 
  group_by(state, type) %>% 
  mutate(cum_value = cumsum(value)) %>% 
  ungroup()

## expand_grid to add zeroes
### use y_o_s for index
state<-unique(y_o_s$statename)
year<-unique(hs$year)
type<-unique(hs$type)
temp<-expand_grid(state, year, type)

hs<-temp %>% 
  left_join(hs) %>% 
  replace_na(list(value = 0,
                  cum_value = 0))

hs <- hs %>% 
  select(-value) %>% 
  pivot_wider(names_from = type,
              values_from = cum_value,
              names_prefix = "hs_")

### add ID and pop vars back to hs

y_o_s <- read_csv("./data/year_of_statehood.csv") %>% 
  left_join(dat %>% 
              select(state, region) %>% 
              distinct()) %>% 
  filter(!(is.na(region)))

hs<-hs %>% 
  rename(statename = state) %>% 
  left_join(y_o_s)

#### join cumulative acres / claims onto dat for the valid census years, then carried forward

dat<-dat %>% 
  left_join(hs)
```

@@

#### Measuring anti-Blackness and post-Reconstruction racial politics

1) POP COMP (BLACK) - VERY CLEAR

2) jim crow salience

3)

#### Measuring the size, power, and interests of the settler polity in active conflict zones

#### Measuring the magnitude of social problems Progressive reformers would seek to 'fix'

4) URBANIZATION + IMMIGRANT POP - CLEAR-ISH. HOW DOES THIS RELATE TO SETTLER INTERESTS?

# Findings

1) walk through the context variables descriptively
2) walk through the state formation variables descriptively
3) estimate a few multivariate models
4) do I need a DAG + causal ID?

## The distributions of Black populations in the US

@@

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = log(blk))) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

@@

## The spatial distribution of the frontier settler polity

### HOMESTEAD ACT

@@

```{r ts-homestead-claims, fig.cap = "Cumulative Homestead act claims by state and year"}
ggplot(hs %>% 
         filter(year < 1940),
       aes(x = year,
           y = hs_claims,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  geom_text(data = hs %>% 
              filter(year == 1935),
            aes(x = year,
                y = hs_claims,
                label = state),
            position = position_nudge(x = 8),
            alpha = 0.7) + 
facet_wrap(~region)  + 
  labs(y = "Number of claims",
       x = "Year",
       subtitle = "") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

### ALTERNATIVE?
# 
# ggplot(hs %>% 
#          filter(year < 1940),
#        aes(x = year,
#            y = hs_claims,
#            group = state,
#            color = region)) + 
#   geom_line(alpha = 0.3) + 
#   geom_text(data = hs %>%
#               filter(year == 1940,
#                      hs_claims > 10000),
#             aes(x = year,
#                 y = hs_claims,
#                 label = state),
#             position = position_nudge(x = 2),
#             size = 3) +
#   geom_line(data = hs %>% 
#               group_by(year, region) %>%
#               filter(year < 1940) %>% 
#               summarize(hs_claims = mean(hs_claims)),
#             aes(x = year, y = hs_claims, color = region,
#                 group = region), linewidth = 2, lty = 2, alpha = 0.7) +
#   labs(y = "Number of claims",
#        x = "Year",
#        subtitle = "Panel 1.") + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

@@

```{r}
dat %>% 
  filter(year == 1940) %>% 
  mutate(region = factor(region, 
                         levels = c("West", 
                                    "North Central",
                                    "South",
                                    "Northeast"))) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  ggplot(aes(x = hs_pop,
             y = fct_reorder(state, hs_pop))) + 
  geom_point() + 
  labs(x = "Cumulative homestead claims per state population, 1940",
       y = "",
       color = "Census Region") + 
  facet_wrap(~region, 
             scales = "free_y")
```

```{r}
dat %>% 
  filter(year == 1940) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = hs_pop)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

### white immigrants and cities

WRITE IT UP

@@

# maps of focal measures

@@

```{r emp-ts-plot, fig.cap="Employment rates per 1000 population 1900 - 1930 (US States, PUMS data)"}
dat %>% 
  filter(year > 1880) %>% 
  group_by(state, year) %>% 
  summarize(police = sum(police) / sum(tot_pop) * 1e3,
            teachers = sum(teachers) / sum(tot_pop) * 1e3,
            socworkers = sum(socwork) / sum(tot_pop) * 1e3) %>% 
  pivot_longer(cols = police:socworkers,
               names_to = "name",
               values_to = "employment")  %>%  
  filter(year < 1940) %>% 
  ggplot(aes(x = year, y = employment)) + 
  geom_line(aes(group = state), alpha = 0.2) + 
  geom_line(data = dat %>% ### add national line
              group_by(year, region) %>%
              filter(year > 1880, year < 1940) %>% 
              summarize(police = sum(police, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3,
                        teachers = sum(teachers, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3,
                        socworkers = sum(socwork, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3) %>%
              pivot_longer(cols = police:socworkers,
                           names_to = "name",
                           values_to = "employment"),
  aes(x = year, y = employment, color = region), linewidth = 1.2) +
  facet_wrap(~name) + 
  scale_y_log10() + 
  labs(subtitle = "note variable y axis scales, national average dashed")
```

@@

dat %>% 
  filter(!(is.na(region))) %>% 
  group_by(region, year) %>% 
  summarize(police = sum(police) / sum(tot_pop) * 1e3,
            teachers = sum(teachers) / sum(tot_pop) * 1e3,
            socworkers = sum(socwork) / sum(tot_pop) * 1e3) %>% 
  pivot_longer(cols = police:socworkers,
               names_to = "name",
               values_to = "employment")  %>%  
  ggplot(aes(x = year, y = employment, 
             color = region)) + 
  geom_line() + 
  facet_wrap(~name, scales = "free", ncol = 3) + 
  labs(subtitle = "note variable y axis scales")
```

@@

### agents data eda

```{r}
agents<-read_csv("./data/AGENTS_DATA_COLLECTION.csv")

agents1895<-agents %>% 
  filter(YEAR == 1895)

state_counts95<-agents1895 %>% 
  group_by(`STATE/SUPERINTENDENCY`) %>% 
  summarize(enrolled = sum(STUDENTS_ENROLL, na.rm=T))
```

NEED A STATE MAPPING ST SOME POINT
I WANT TO GET STUDENT POPS BY STATE...

USE BIA DATA FOR NOW FOR RESIDENTIAL / NON

```{r}
bs<-read_csv("./data/Boarding_Schools.csv")

bs <- bs %>% 
  group_by(State, ST_Reservation) %>% 
  summarize(n = n()) %>% 
  filter(!(is.na(ST_Reservation)))
```

>>>>

---
title: "Varying paths to building the social state in the progressive era"
author: "Frank Edwards"
date: "last updated 11/1/23"
output: bookdown::html_document2
---

@@

# TO DO

1. pull immig + urban data
2. pull tenant farmer data
3. model employment -> each, consider loglinear process?

theory re: white immigs: coercion / rehab in NE, yeoman uplift in W, frontier as panacea for urban, lots of hist on that, gets at SocWork/Police vs Teachers employment

@@

```{r read, cache = T}
# ### on libra, read raw
# ### read employment data
# ### full ts of occupations ADD CONSTABLES, THERES A DECENT NUMBER OF THEM
# dat_occ<-read_csv("./data/usa_00071.csv") %>% 
#   rename_all(tolower) %>% 
#   mutate(occ1950 = 
#            case_when(
#              occ1950 == 79 ~ "socwork",
#              occ1950 == 773 ~ "police",
#              occ1950 == 771 ~ "police",
#              occ1950 == 782 ~ "police",
#              occ1950 == 93 ~ "teachers"
#            )) %>% 
#   group_by(year, statefip, occ1950) %>% 
#   summarize(emp = sum(perwt)) %>% 
#   pivot_wider(names_from = occ1950, values_from = emp,
#               values_fill = 0)
# ### AK OR HI HERE? check that at IPUMS
# 
# # pop data (total, age, race, nativitiy, urban)
# dat_immig<-read_csv("./data/usa_00070.csv") %>% 
#   rename_all(tolower) %>% 
#   group_by(year, statefip) %>% 
#   summarize(tot_pop = sum(perwt),
#             children = sum((age<18) * perwt),
#             white_immig = 
#               sum((nativity == 5) * (race == 1) * (hispan == 0) * perwt),
#             blk = 
#               sum((race == 2) * perwt),
#             urban = sum(hhwt * (urban ==2)),
#             tot_hh = sum(hhwt)) 
# ### AK OR HI HERE? check that at IPUMS
# 
# 
# ### group quarters data, 100% 
# gq<-read_csv("./data/usa_00069.csv") %>% 
#   rename_all(tolower) %>% 
#   filter(age<18) %>% 
#   group_by(year, statefip, gqtype) %>% 
#   summarize(gq_pop = sum(perwt, na.rm = T)) %>% 
#   mutate(gqtype = case_when(
#     gqtype == 2 ~ "child_inst_correctional",
#     gqtype == 3 ~ "child_inst_mental",
#     gqtype == 4 ~ "child_inst_poor_disabled")) %>% 
#   filter(gqtype!="other") %>% 
#   pivot_wider(names_from = gqtype, values_from = gq_pop,
#               values_fill = 0)
# ### AK OR HI HERE? check that at IPUMS
# 
#  
# ### join
# dat <- data.frame(state = state.abb, 
#                region = state.region, 
#                division = state.division) %>% 
#   left_join(state.fips %>% 
#               select(fips, abb) %>% 
#               distinct() %>% 
#               rename(state = abb, statefip = fips)) %>% 
#   left_join(dat_immig) %>% 
#   left_join(dat_occ) %>% 
#   ungroup() %>% 
#   left_join(gq)
# 
# ### homestead act data
# #### load in homestead data and boarding school data
# ### state measure can be cumulative sum of homesteads, that gets at polity
# ### homesteading data
# y_o_s <- read_csv("./data/year_of_statehood.csv") %>% 
#   left_join(dat %>% 
#               select(state, region) %>% 
#               distinct()) 
# 
# hs<-read_csv("./data/blm_homesteads_clean.csv") %>% 
#   pivot_longer(claims1868:acres1961,
#                names_to = "name",
#                values_to = "value") %>% 
#   mutate(year = str_sub(
#     name, start = -4, end = -1),
#     type = str_sub(
#       name, start = 1, end = -5)) %>% 
#   select(-name) %>% 
#   mutate(year = as.numeric(year)) %>% 
#   mutate(value = ifelse(is.na(value),
#                         0,
#                         value))
# 
# ### split Dakota Territory values to 0.5 ND, 0.5 SD to preserve time series
# hs_dakota<-hs %>% 
#   filter(state == "Dakota Territory") %>% 
#   mutate(value = value / 2) %>% 
#   select(-state)
# 
# hs<-hs %>% 
#   bind_rows(hs_dakota %>% 
#               mutate(state = "North Dakota")) %>% 
#   bind_rows(hs_dakota %>% 
#               mutate( state = "South Dakota")) %>% 
#   filter(state != "Dakota Territory") %>% 
#   group_by(state, year, type) %>% 
#   summarize(value = sum(value))
#   
# hs<-hs %>% 
#   arrange(state, year) %>% 
#   group_by(state, type) %>% 
#   mutate(cum_value = cumsum(value)) %>% 
#   ungroup()
# 
# ## expand_grid to add zeroes
# ### use y_o_s for index
# state<-unique(y_o_s$statename)
# year<-unique(hs$year)
# type<-unique(hs$type)
# temp<-expand_grid(state, year, type)
# 
# hs<-temp %>% 
#   left_join(hs) %>% 
#   replace_na(list(value = 0,
#                   cum_value = 0))
# 
# hs <- hs %>% 
#   select(-value) %>% 
#   pivot_wider(names_from = type,
#               values_from = cum_value,
#               names_prefix = "hs_") 
# 
# ### add ID and pop vars back to hs
# 
# y_o_s <- read_csv("./data/year_of_statehood.csv") %>% 
#   left_join(dat %>% 
#               select(state, region) %>% 
#               distinct()) %>% 
#   filter(!(is.na(region)))
# 
# hs<-hs %>% 
#   rename(statename = state) %>% 
#   left_join(y_o_s)
# 
# #### join cumulative acres / claims onto dat for the valid census years, then carried forward
# 
# dat<-dat %>% 
#   left_join(hs)

# on local read temp
dat<-read_csv("./data/temp_data.csv")
hs<-read_csv("./data/temp_hs.csv")
```

@@

### Measuring anti-Blackness and post-Reconstruction racial politics

1) jim crow salience (can we define as brutal coupling of formal / informal? less state capacity built for maintenance / strategic underinvestment)
2) exclusion / formal soc control salience in urban non-S

POPULATION ALONE DESCRIPTIVES

@@

@@

```{r}
dat %>% 
  filter(year == 1930) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = log(blk))) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

### Measuring the size, power, and interests of the expropriative settler polity

@@

```{r ts-homestead-claims, fig.cap = "Cumulative Homestead act claims by state and year"}
ggplot(hs %>% 
         filter(year < 1940),
       aes(x = year,
           y = hs_claims,
           group = state)) + 
  geom_line(alpha = 0.5) + 
  geom_text(data = hs %>% 
              filter(year == 1935),
            aes(x = year,
                y = hs_claims,
                label = state),
            position = position_nudge(x = 8),
            alpha = 0.7) + 
facet_wrap(~region)  + 
  labs(y = "Number of claims",
       x = "Year",
       subtitle = "") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(region = factor(region, 
                         levels = c("West", 
                                    "North Central",
                                    "South",
                                    "Northeast"))) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  ggplot(aes(x = hs_pop,
             y = fct_reorder(state, hs_pop))) + 
  geom_point() + 
  labs(x = "Cumulative homestead claims per state population, 1920",
       y = "",
       color = "Census Region") + 
  facet_wrap(~region, 
             scales = "free_y")
```

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(hs_pop = hs_claims / tot_pop) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = hs_pop)) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

### Measuring the magnitude of social problems Progressive reformers would seek to 'fix'

PULL IMMIG + URBAN - UNIQUE COMBO THAT PUSHES FORWARD WELFARE STATE

@@

```{r}
dat %>% 
  filter(year == 1920) %>% 
  mutate(state = tolower(statename)) %>% 
  ggplot(aes(fill = log(white_immig))) + 
  geom_map(aes(map_id = state), map = map) + 
  expand_limits(x = map$long, y = map$lat) + 
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  theme(legend.position = "bottom")
```

@@

```{r emp-ts-plot, fig.cap="Employment rates per 1000 population 1900 - 1930 (US States, PUMS data)"}
dat %>% 
  filter(year > 1880) %>% 
  group_by(state, year) %>% 
  summarize(police = sum(police) / sum(tot_pop) * 1e3,
            teachers = sum(teachers) / sum(tot_pop) * 1e3,
            socworkers = sum(socwork) / sum(tot_pop) * 1e3) %>% 
  pivot_longer(cols = police:socworkers,
               names_to = "name",
               values_to = "employment")  %>%  
  filter(year < 1940) %>% 
  ggplot(aes(x = year, y = employment)) + 
  geom_line(aes(group = state), alpha = 0.1) + 
  geom_line(data = dat %>% ### add national line
              group_by(year, region) %>%
              filter(year > 1880, year < 1940) %>% 
              summarize(police = sum(police, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3,
                        teachers = sum(teachers, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3,
                        socworkers = sum(socwork, na.rm=T) / sum(tot_pop, na.rm=T) * 1e3) %>%
              pivot_longer(cols = police:socworkers,
                           names_to = "name",
                           values_to = "employment"),
  aes(x = year, y = employment, color = region), linewidth = 1) +
  facet_wrap(~name) + 
  scale_y_log10() + 
  labs(subtitle = "note variable y axis scales, national average dashed",
       y = "Employment per 1,000 population (note log scale)") + 
  theme(legend.position = "bottom")
```

@@

dat %>% 
  filter(!(is.na(region))) %>% 
  group_by(region, year) %>% 
  summarize(police = sum(police) / sum(tot_pop) * 1e3,
            teachers = sum(teachers) / sum(tot_pop) * 1e3,
            socworkers = sum(socwork) / sum(tot_pop) * 1e3) %>% 
  pivot_longer(cols = police:socworkers,
               names_to = "name",
               values_to = "employment")  %>%  
  ggplot(aes(x = year, y = employment, 
             color = region)) + 
  geom_line() + 
  scale_y_log10() + 
  facet_wrap(~name, ncol = 3) + 
  labs(subtitle = "note variable y axis scales")
```

@@

## PRELIMINARY MODELS

employment trajectories look kinda log-linear. almost certainly non-linear, so need a curve on trajectories regardless. log probably solves it well enough.

q1: salience of settler expropriotion -> state formation

four models, each employment outcome + children in poverty institutions

start with lm for log(rates)

trouble fitting models for sparse data (e.g. social workers, children in gq), but works fine and in expected directions for teachers, police.

switch to brms for estimation on tricky ones

```{r}
### FIT PROBLEM WAS WITH EARLY YEARS NEAR ZERO AND LOW VARIANCE
model_dat <- dat %>% 
  filter(year>=1880) %>% 
  mutate(police_rate = (police + 1) / tot_pop * 1e5,
         socwork_rate = (socwork + 1) / tot_pop * 1e5,
         teacher_rate = (teachers + 1) / tot_pop * 1e5,
         child_inst_rate = (child_inst_poor_disabled + 0.1) / 
           children * 1e5,
         hsclaim_prop = hs_claims / tot_pop,
         year = scale(year))

sc_police<-lmer(log(police_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

sc_teacher<-lmer(log(teacher_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

sc_socwork<-brm(log(socwork_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat)

sc_gq<-brm(log(child_inst_rate) ~ 
              scale(hsclaim_prop) +
              year + 
              (1 + year|state),
            data = model_dat)
```

This all looks fine. How do I want to visualize this?

we'll eventually do trajectory visuals for each regime type - what is the expected trajectory under different regimes, where do we end up in 1930? Think about how to handle to intercepts / slopes. What would a zero mean? pulling out everything not directly in national trend and regime measures. is that what I'm after? I guess I really do just want the regime-specific trend lines, in which case E[rate | regime] should zero out the individual idioscyncracies. it's oversimplified, but that should be ok. Just be super clear on what the iterpretation is on prediction. it's not a hypothetical state, its the mean expectation for the slope of the regime var.

model is

\[\hat{Y} = \alpha + \beta \textrm{Regime} + \delta year + \epsilon\]

Here we are very much interested just in $\beta$, the rest are nuissance parameters. that works. can vis e[state type] under diffferent REGIME sets.

Q: are there places with competing regime sets? CA, WA, others? where we need a multivariate regime model?

```{r}
new_dat<-data.frame(year =  c(1.3, 1.3, 1.3),
                    hsclaim_prop = c(0, 0.05, 0.15))
t<-predict(sc_police, 
           newdata = new_dat,
           re.form = ~0,
           type = "link")

t1<-predict(sc_teacher, 
           newdata = new_dat,
           re.form = ~0,
           type = "link")

exp(t1)
```
